---
title: "Cue analysis"
author: "Liam D. Bailey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CueAnalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}

devtools::load_all()
devtools::install_github("LiamDBailey/MyFuncs", upgrade = "never")

#Load libraries
library(MyFuncs)

```

Load the settlement data

```{r}

data("SETTLE_poisson")

```

Now we know that settlement patterns have changed over time (i.e. elevation doesn't affect settlement in the same way in later years).

Next, we want to consider what cues might be used to inform settlement behaviour.

We consider 4 potential cues:
- Direct elevation cue
- Food availability cue (i.e. coast/gully distance)
- Conspecific attraction (territory density in previous year)
- Conspecific reproductive experience (Relative fledgling success in the area in previous year)

```{r}

SETTLE_poisson$Median3 <- transform(x = SETTLE_poisson$Median2, type = "standard.log")
SETTLE_poisson$Gully_Dist2 <- transform(x = SETTLE_poisson$Gully_Dist, type = "standard.log")
SETTLE_poisson$Coast_Dist2  <- transform(x = SETTLE_poisson$Coast_Dist, type = "standard.log")
SETTLE_poisson$Grid_area2   <- transform(x = SETTLE_poisson$Grid_area, type = "standard")
SETTLE_poisson$Year_group <- SETTLE_poisson$Year - min(SETTLE_poisson$Year)

```

We are going to consider density and fledgling data over the past 3 years (rather than just 1). This is a compromise between including more years and keeping similar sample size.

```{r}

SETTLE_poisson <- filter(SETTLE_poisson, !is.na(rel_density3yr))

```

```{r}

full_mod <- spaMM::fitme(lgl_Settle ~ Year_group * log(Median3) + log(Coast_Dist2) + log(Gully_Dist2) + rel_fledge_est3yr + rel_density3yr + Grid_area2 + (1|Sub_Area),
                  data = SETTLE_poisson, family = binomial(logit), method = "PQL/L")

```

Assess linearity of relationships

```{r, fig.width = 7, fig.height = 5}

SETTLE_poisson %>%
  ungroup() %>%
  mutate(logits = c(predict(full_mod)), Median2 = log(Median2)) %>%
  dplyr::select(logits, Year, Median2, rel_fledge_est3yr, rel_density3yr) %>%
  tidyr::gather(key = "Predictor", value = "Pred_value", -logits) %>%
  ggplot()+
  geom_point(aes(x = Pred_value, y = logits))+
  facet_wrap(facets = ~Predictor, scales = "free_x")

```

Test residuals using DHARMa package

```{r, message = FALSE, fig.width = 7, fig.height = 5}

set.seed(666)

simResiduals <- DHARMa::simulateResiduals(fittedModel = full_mod, n = 5000)

DHARMa::plotSimulatedResiduals(simulationOutput = simResiduals)

```

Test VIF

```{r}

car::vif(glm(lgl_Settle ~ Year_group + log(Median3) + log(Coast_Dist2) + log(Gully_Dist2) + Grid_area2 + rel_fledge_est3yr + rel_density3yr, family = "binomial", data = SETTLE_poisson))

```

No real issues.

---------------------------

Test residual of random effects

```{r}

qqnorm(spaMM::ranef(full_mod)[[1]])
qqline(spaMM::ranef(full_mod)[[1]])

```

Sub area is fine.

-----------------------

To compare our different hypothesese (i.e. different cues) we will use AIC model selection.

Model selection:

```{r}

#Fit models with every possible combination of cues (using both spamm and lme4 for R2 values)
all_models <- expand.grid(elev = c("", "+ Year_group : log(Median3)"),
                          density = c("", "+ rel_density3yr"),
                          food = c("", "+ log(Coast_Dist2) + log(Gully_Dist2)"),
                          fledge = c("", "+ rel_fledge_est3yr")) %>% 
  as_tibble() %>% 
  mutate(mod_formula = glue::glue("lgl_Settle ~ Year_group + log(Median3) + Grid_area2 + (1|Sub_Area) {cues}",
                                             cues = paste0(elev, density, food, fledge))) %>% 
  rowwise() %>% 
  mutate(models_spamm = list(spaMM::fitme(as.formula(mod_formula), data = SETTLE_poisson, family = binomial(logit), method = "PQL/L")),
         models_lme4 = list(lme4::glmer(as.formula(mod_formula), data = SETTLE_poisson, family = binomial(logit))))

#This null model with lme4 is needed to improve fitting for R2
#Otherwise, the null model is re-estimated every time.
null_model <- lme4::glmer(lgl_Settle ~ 1 + (1|Sub_Area),
                    data = SETTLE_poisson, family = binomial(logit))

#Determine AIC values for every model.
(selection_table <- all_models %>% 
  mutate(cAIC = as.numeric(AIC(models_spamm, verbose = FALSE)[2]),
         R2s = list(MuMIn::r.squaredGLMM(models_lme4, null = null_model)),
         R2m = R2s[1, "R2m"],
         R2c = R2s[1, "R2c"],
         elev = if(elev != "") "+" else "",
         food = if(food != "") "+" else "",
         fledge = if(fledge != "") "+" else "",
         density = if(density != "") "+" else "") %>%
  ungroup() %>% 
  arrange(cAIC) %>% 
  mutate(deltaAIC = cAIC - min(cAIC),
         relative_loglik = exp(-0.5*deltaAIC),
         AIC_weight = relative_loglik/sum(relative_loglik),
         Conf_set = cumsum(AIC_weight)) %>% 
  dplyr::select(elev:fledge, cAIC, deltaAIC:Conf_set, R2m, R2c, models_spamm, models_lme4, -mod_formula, -relative_loglik, -R2s))

```

We're cheating a bit here by fitting with a different model package to estimate R2. However, the coefficient estimates and AIC values are essentially identical between models.

We need to compare marginal AIC because conditional AIC is only available with spaMM

```{r}

selection_table %>% 
  rowwise() %>% 
  mutate(AIC_spamm = as.numeric(AIC(models_spamm, verbose = FALSE)[1]),
         AIC_lme4 = as.numeric(AIC(models_lme4)),
         diff = abs(AIC_spamm - AIC_lme4)) %>% 
  pull(diff) %>% 
  range

```

Difference in AIC estimation is >=0.02 units. So we can feel pretty confident that we're achieving very similar fits with the two modelling approaches.

Print selection table with kableExtra::kable

```{r}

selection_table %>% 
  dplyr::select(elev:R2c) %>%
  rename(Elevation = elev,
         Food = food,
         Fledglings = fledge,
         Density = density,
         `Cumulative weight` = Conf_set,
         wi = AIC_weight) %>% 
  mutate_at(vars(cAIC, deltaAIC, R2m, R2c), ~format(round(., digits = 2))) %>%
  rowwise() %>% 
  mutate_at(vars(wi), ~if(. < 0.01) "<0.01" else format(round(., digits = 2))) %>%
  mutate_at(vars(`Cumulative weight`), ~if(. > 0.9949) ">0.99" else format(round(., digits = 2))) %>%
  ungroup() %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "bordered")

```

-----------------------

Plot effect of conspecific density

```{r, fig.width = 7, fig.height = 5}

plot_dat <- expand.grid(Year_group = mean(SETTLE_poisson$Median3),
                        Median3 = mean(SETTLE_poisson$Median3),
                        Coast_Dist2 = mean(SETTLE_poisson$Coast_Dist2),
                        Gully_Dist2 = mean(SETTLE_poisson$Gully_Dist2),
                        rel_density3yr = seq(min(SETTLE_poisson$rel_density3yr),
                                          max(SETTLE_poisson$rel_density3yr),
                                          length.out = 1000),
                        rel_fledge_est3yr = mean(SETTLE_poisson$rel_fledge_est3yr),
                        Grid_area2 = mean(SETTLE_poisson$Grid_area2))

predictions <- predict(selection_table$models_spamm[[1]], newdata = plot_dat, re.form = NA,
                       intervals = "predVar")

plot_dat <- plot_dat %>% 
  mutate(pred = as.numeric(predictions[, 1]),
         lower = as.numeric(attr(predictions, "intervals")[, 1]),
         upper = as.numeric(attr(predictions, "intervals")[, 2]))

bin_dat <- SETTLE_poisson %>%
  ungroup() %>% 
  mutate(Density_class = cut(.$rel_density3yr, breaks = c(seq(min(SETTLE_poisson$rel_density3yr), max(SETTLE_poisson$rel_density3yr), length.out = 20)), right = FALSE, include.lowest = TRUE)) %>% 
  group_by(Density_class) %>% 
  summarise(mean_settle = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[4]),
            lower = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[5]),
            upper = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[6]),
            n = n()) %>% 
  filter(!is.na(mean_settle)) %>%
  mutate(Density_numeric = purrr::pmap_dbl(.l = list(Density_class = as.character(.$Density_class)),
                                  .f = function(Density_class){
                                    
                                    return(mean(as.numeric(strsplit(split = ",", x = gsub("\\(|\\]|\\[|\\)", "", Density_class))[[1]])))
                                    
                                  }))
  

(density <- ggplot()+
  geom_ribbon(data = plot_dat, aes(x = rel_density3yr, ymin = lower, ymax = upper), fill = "dark grey", colour = NA, alpha = 0.25)+
  geom_path(data = plot_dat, aes(x = rel_density3yr, y = pred), lty = 2, size = 1)+
  geom_errorbar(data = bin_dat, aes(x = Density_numeric, ymin = lower, ymax = upper), width = 0.1, size = 0.75)+
  geom_point(data = bin_dat, aes(x = Density_numeric, y = mean_settle), size = 3, shape = 21, fill = "white", stroke = 1)+
  geom_text(data = bin_dat, aes(x = Density_numeric, y = lower - 0.03, label = n), size = 3, family = "sans")+
  theme_ubuntu(base_size = 10)+
  theme(text = element_text(family = "sans"))+
  xlab("Relative conspecific density/0.25 ha grid")+
  ylab("Probability of settlement"))

```

Plot effect of conspecific density

```{r, fig.width = 7, fig.height = 5}

plot_dat <- expand.grid(Year_group = mean(SETTLE_poisson$Median3),
                        Median3 = mean(SETTLE_poisson$Median3),
                        Coast_Dist2 = mean(SETTLE_poisson$Coast_Dist2),
                        Gully_Dist2 = mean(SETTLE_poisson$Gully_Dist2),
                        rel_fledge_est3yr = seq(min(SETTLE_poisson$rel_fledge_est3yr),
                                          max(SETTLE_poisson$rel_fledge_est3yr),
                                          length.out = 1000),
                        rel_density3yr = mean(SETTLE_poisson$rel_density3yr),
                        Grid_area2 = mean(SETTLE_poisson$Grid_area2))

predictions <- predict(selection_table$models_spamm[[1]], newdata = plot_dat, re.form = NA,
                       intervals = "predVar")

plot_dat <- plot_dat %>% 
  mutate(pred = as.numeric(predictions[, 1]),
         lower = as.numeric(attr(predictions, "intervals")[, 1]),
         upper = as.numeric(attr(predictions, "intervals")[, 2]))

bin_dat <- SETTLE_poisson %>%
  ungroup() %>% 
  mutate(Density_class = cut(.$rel_fledge_est3yr, breaks = c(seq(min(SETTLE_poisson$rel_fledge_est3yr), 1, length.out = 20), 1.2, 1.4, 1.6), right = FALSE, include.lowest = TRUE)) %>% 
  group_by(Density_class) %>% 
  summarise(mean_settle = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[4]),
            lower = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[5]),
            upper = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[6]),
            n = n()) %>% 
  filter(!is.na(mean_settle)) %>%
  mutate(Density_numeric = purrr::pmap_dbl(.l = list(Density_class = as.character(.$Density_class)),
                                  .f = function(Density_class){
                                    
                                    return(mean(as.numeric(strsplit(split = ",", x = gsub("\\(|\\]|\\[|\\)", "", Density_class))[[1]])))
                                    
                                  }))
  

(fledge <- ggplot()+
  geom_ribbon(data = plot_dat, aes(x = rel_fledge_est3yr, ymin = lower, ymax = upper), fill = "dark grey", colour = NA, alpha = 0.25)+
  geom_path(data = plot_dat, aes(x = rel_fledge_est3yr, y = pred), lty = 2, size = 1)+
  geom_errorbar(data = bin_dat, aes(x = Density_numeric, ymin = lower, ymax = upper), width = 0.035, size = 0.75)+
  geom_point(data = bin_dat, aes(x = Density_numeric, y = mean_settle), size = 3, shape = 21, fill = "white", stroke = 1)+
  geom_text(data = bin_dat, aes(x = Density_numeric, y = lower - 0.03, label = n), size = 3, family = "sans")+
  scale_y_continuous(limits = c(-0.05, 1), breaks = c(0, 0.25, 0.5, 0.75, 1))+
  theme_ubuntu(base_size = 10)+
  theme(text = element_text(family = "sans"))+
  xlab("Relative conspecific fledgling output/0.25 ha grid")+
  ylab("Probability of settlement"))

```

```{r, fig.width = 14, fig.height = 5}

cowplot::plot_grid(density, fledge, nrow = 1, labels = c("a)", "b)"))

```

```{r}

#Save as pdf
ggsave("../plots/Figure_4.pdf", width = 34, height = 15, units = "cm", dpi = 600)

```

```{r}

CIs <- as.data.frame(cbind(confint(selection_table$models_lme4[[1]], method = "Wald")[-1, ],
             confint(selection_table$models_lme4[[1]], level = 0.995, method = "Wald")[-1, ])) %>%
  rowwise() %>% 
  mutate(CI95 = paste(format(`2.5 %`, digits = 2, scientific = -2, nsmall = 2), format(`97.5 %`, digits = 2, scientific = -2, nsmall = 2), sep = "/"),
         CI995 = paste(format(`0.25 %`, digits = 2, scientific = -2, nsmall = 2), format(`99.75 %`, digits = 2, scientific = -2, nsmall = 2), sep = "/")) %>% 
  ungroup() %>%
  mutate(Estimate = format(round((spaMM::fixef(selection_table$models_lme4[[1]])), digits = 2), nsmall = 2),
         Variable = names(spaMM::fixef(selection_table$models_lme4[[1]]))) %>% 
  dplyr::select(Variable, Estimate, CI95, CI995)

(CI_25ha <- CIs %>% 
  dplyr::select(Variable, Estimate, CI95, CI995) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "bordered"))

```


----------------------------------

##100x100

Load the settlement data

```{r}

data("SETTLE_poisson_100")

```

Now we know that settlement patterns have changed over time (i.e. elevation doesn't affect settlement in the same way in later years).

Next, we want to consider what cues might be used to inform settlement behaviour.

We consider:

- Median elevation of the area and coastal distance (i.e. elevation cues)
- Territory density in the year BEFORE settlement (i.e. conspecific attraction)
- Fledgling production in the year BEFORE settlement (i.e. reproductive information)

```{r}

SETTLE_poisson_100$Median3 <- transform(x = SETTLE_poisson_100$Median2, type = "standard.log")
SETTLE_poisson_100$Gully_Dist2 <- transform(x = SETTLE_poisson_100$Gully_Dist, type = "standard.log")
SETTLE_poisson_100$Coast_Dist2  <- transform(x = SETTLE_poisson_100$Coast_Dist, type = "standard.log")
SETTLE_poisson_100$Grid_area2   <- transform(x = SETTLE_poisson_100$Grid_area, type = "standard")
SETTLE_poisson_100$Year_group <- SETTLE_poisson_100$Year - min(SETTLE_poisson_100$Year)

```

```{r}

SETTLE_poisson_100 <- filter(SETTLE_poisson_100, !is.na(rel_density3yr))

```

```{r}

full_mod_100 <- spaMM::fitme(lgl_Settle ~ Year_group * log(Median3) + rel_fledge_est3yr + rel_density3yr + Grid_area2 + (1|Sub_Area),
                  data = SETTLE_poisson_100, family = binomial(logit), method = "PQL/L")

```

Assess linearity of relationships

```{r, fig.width = 7, fig.height = 5}

SETTLE_poisson_100 %>%
  ungroup() %>%
  mutate(logits = c(predict(full_mod_100)), Median2 = log(Median3)) %>%
  dplyr::select(logits, Year, Median2, rel_fledge_est3yr, rel_density3yr) %>%
  tidyr::gather(key = "Predictor", value = "Pred_value", -logits) %>%
  ggplot()+
  geom_point(aes(x = Pred_value, y = logits))+
  facet_wrap(facets = ~Predictor, scales = "free_x")

```

```{r}

summary(full_mod_100)

```

Test residuals using DHARMa package

```{r, message = FALSE, fig.width = 7, fig.height = 5}

set.seed(666)

simResiduals <- DHARMa::simulateResiduals(fittedModel = full_mod_100, n = 5000)

DHARMa::plotSimulatedResiduals(simulationOutput = simResiduals)

```

Test VIF

```{r}

car::vif(glm(lgl_Settle ~ Year_group + log(Median3) + log(Coast_Dist2) + log(Gully_Dist2) + rel_fledge_est3yr + Grid_area2 + rel_density3yr, family = "binomial", data = SETTLE_poisson_100))

```

No real issues.

---------------------------

Test residual of random effects

```{r}

qqnorm(spaMM::ranef(full_mod_100)[[1]])
qqline(spaMM::ranef(full_mod_100)[[1]])

```

Sub area is fine.

-----------------------

Plot effect of conspecific density

```{r, fig.width = 7, fig.height = 5}

plot_dat <- expand.grid(Year_group = mean(SETTLE_poisson_100$Year_group),
                        Median3 = mean(SETTLE_poisson_100$Median3),
                        Coast_Dist2 = mean(SETTLE_poisson_100$Coast_Dist2),
                        Gully_Dist2 = mean(SETTLE_poisson_100$Gully_Dist2),
                        rel_density3yr = seq(min(SETTLE_poisson_100$rel_density3yr, na.rm = T),
                                          max(SETTLE_poisson_100$rel_density3yr, na.rm = T),
                                          length.out = 1000),
                        rel_fledge_est3yr = mean(SETTLE_poisson_100$rel_fledge_est3yr, na.rm = T),
                        Grid_area2 = mean(SETTLE_poisson_100$Grid_area2))

predictions <- predict(full_mod_100, newdata = plot_dat, re.form = NA,
                       intervals = "predVar")

plot_dat <- plot_dat %>% 
  mutate(pred = as.numeric(predictions[, 1]),
         lower = as.numeric(attr(predictions, "intervals")[, 1]),
         upper = as.numeric(attr(predictions, "intervals")[, 2]))

bin_dat <- SETTLE_poisson_100 %>%
  ungroup() %>% 
  mutate(Density_class = cut(.$rel_density3yr, breaks = c(seq(min(SETTLE_poisson_100$rel_density3yr, na.rm = T), max(SETTLE_poisson_100$rel_density3yr, na.rm = T), length.out = 20)), right = FALSE, include.lowest = TRUE)) %>% 
  group_by(Density_class) %>% 
  summarise(mean_settle = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[4]),
            lower = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[5]),
            upper = as.numeric(binom::binom.wilson(x = sum(lgl_Settle), n = n())[6]),
            n = n()) %>% 
  filter(!is.na(mean_settle)) %>% 
  mutate(Density_numeric = purrr::pmap_dbl(.l = list(Density_class = as.character(.$Density_class)),
                                  .f = function(Density_class){
                                    
                                    return(mean(as.numeric(strsplit(split = ",", x = gsub("\\(|\\]|\\[|\\)", "", Density_class))[[1]])))
                                    
                                  }))
  

ggplot()+
  geom_ribbon(data = plot_dat, aes(x = rel_density3yr, ymin = lower, ymax = upper), fill = "dark grey", colour = NA, alpha = 0.25)+
  geom_path(data = plot_dat, aes(x = rel_density3yr, y = pred), lty = 2, size = 1)+
  geom_errorbar(data = bin_dat, aes(x = Density_numeric, ymin = lower, ymax = upper), width = 0.1, size = 1)+
  geom_point(data = bin_dat, aes(x = Density_numeric, y = mean_settle), size = 3, shape = 21, fill = "white", stroke = 1)+
  geom_text(data = bin_dat, aes(x = Density_numeric, y = lower - 0.03, label = n), size = 3, family = "Ubuntu")+
  theme_ubuntu()+
  xlab("Relative territory density/ha grid")+
  ylab("Probability of settlement")

```

```{r}

CI_mod <- lme4::glmer(lgl_Settle ~ Year_group * log(Median3) + Grid_area2 + rel_density3yr + rel_fledge_est3yr + (1|Sub_Area),
                  data = SETTLE_poisson_100, family = binomial(logit))

CIs <- as.data.frame(cbind(confint(CI_mod, method = "Wald")[-1, ],
             confint(CI_mod, level = 0.995, method = "Wald")[-1, ])) %>%
  rowwise() %>% 
  mutate(CI95 = paste(format(`2.5 %`, digits = 2, scientific = -2, nsmall = 2), format(`97.5 %`, digits = 2, scientific = -2, nsmall = 2), sep = "/"),
         CI995 = paste(format(`0.25 %`, digits = 2, scientific = -2, nsmall = 2), format(`99.75 %`, digits = 2, scientific = -2, nsmall = 2), sep = "/")) %>% 
  ungroup() %>%
  mutate(Estimate = format(round((spaMM::fixef(CI_mod)), digits = 2), nsmall = 2),
         Variable = names(spaMM::fixef(CI_mod))) %>% 
  dplyr::select(Variable, Estimate, CI95, CI995)

(CI_100ha <- CIs %>% 
  dplyr::select(Variable, Estimate, CI95, CI995) %>% 
    kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "bordered"))

```

Determine R2 values for reviewers. To do this we need to use our lme4 model.

Add this to our CI estimates.

```{r}

(R2 <- MuMIn::r.squaredGLMM(CI_mod))

```
