---
title: "Change in territory vacancy patterns"
author: "Liam D. Bailey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vacancy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}

devtools::load_all()
devtools::install_github("LiamDBailey/MyFuncs", upgrade = "never")

#Load libraries
library(MyFuncs)

```

Load our data

```{r}

data("VacatedTerritories")

```

We want to determine whether there is an overall relationship between median territory elevation and territory vacancy rates. As before, we will use a mixed effects model with MaleID and Area as random effects.

We considered:

- Median elevation of a territory.
- Year.
- Distance from the coast (this includes information on territory quality, both absolute status and territory quality e.g. 1st/2nd row).
- Distance from the gully.
- Status (To better differentiate between leapfrog and resident behaviour)

We standardised our variables for analysis (sd = 1). This was needed to acheive convergence with variables that have widely different scales. However, we did not centre to a mean of 0 as log transformation was used.  

```{r}

#We want to use centred variables
VacatedTerritories$Median3     <- transform(x = VacatedTerritories$Median2, type = "standard.log")
VacatedTerritories$Year_group  <- VacatedTerritories$Year - min(VacatedTerritories$Year)
VacatedTerritories$Coast_Dist2 <- transform(x = VacatedTerritories$Coast_Dist, type = "standard.log")
VacatedTerritories$Gully_Dist2 <- transform(x = VacatedTerritories$Gully_Dist, type = "standard.log")

```

We considered a logarithmic relationship for elev, coast distance and gully distance as we predict that the benefit of these variables will decline.

Our models are not mutually exclusive and we have no issue with power. Therefore, we are going to just look at the global model.

```{r}

full_mod <- spaMM::fitme(Vacated ~ log(Median3) * Year_group + Status + (1|Sub_Area), family = "binomial", data = VacatedTerritories, method = "PQL/L")

```

The only reason why we wouldn't do this is if there were issues with VIF.

```{r}

car::vif(glm(Vacated ~ log(Median3) + Year_group + Status, family = "binomial", data = VacatedTerritories))

```

Our variables have reasonable VIF levels. Therefore, we can include them with no concern how they will impact our important variables.

1. Test residuals using DHARMa package

```{r, message = FALSE, fig.height=5, fig.width=5}

set.seed(666)

simResiduals <- DHARMa::simulateResiduals(fittedModel = full_mod, n = 5000)

DHARMa::plotSimulatedResiduals(simulationOutput = simResiduals)

```

Seems ok.

With plotly

```{r, fig.width = 7, fig.height = 5}

new_dat <- expand.grid(Median3 = seq(min(VacatedTerritories$Median3),
                                     max(VacatedTerritories$Median3), length.out = 100),
                       Year_group = seq(min(VacatedTerritories$Year_group),
                                        max(VacatedTerritories$Year_group), 1),
                       Grid_area2 = mean(VacatedTerritories$Grid_area2), Coast_Dist2 = 1,
                       Gully_Dist2 = 1, Status = "resident")

new_dat$pred <- as.numeric(predict(full_mod, newdata = new_dat, re.form = NA)[, 1])
new_dat$Median2 <- back.transform(x = new_dat$Median3,
                                  y = VacatedTerritories$Median2,
                                  type = "standard.log")

require(plotly)

(Fig3 <- plot_ly(new_dat, x = ~Year_group, y = ~Median2, z = ~pred, type = "contour", showlegend = F, 
        hoverlabel = list(font =  list(family = "Ubuntu")),
        contours = list(coloring = "fill",
                        showlabels = TRUE,
                        labelfont = list(family = "Ubuntu",
                                         size = 15,
                                         color = "white")),
        colorbar = list(tickfont = list(family = "Ubuntu",
                                        size = 20),
                        title = "Vacancy \n probability",
                        titlefont = list(family = "Ubuntu",
                                         size = 15))) %>%
  layout(xaxis = list(title = "Year",
                  titlefont = list(family = "Ubuntu",
                                   size = 20),
                  tickmode = "array",
                  tickvals = seq(1985, 2015, 5) - min(VacatedTerritories$Year),
                  ticktext = seq(1985, 2015, 5)),
         yaxis = list(title = "Elevation (cm above MHT)",
                      titlefont = list(family = "Ubuntu",
                                   size = 20))))

```

**SAVING AS PDF IS NOT POSSIBLE IN R WITHOUT A PLOTLY PREMIUM SUBSCRICPTION**
**WE EXPORT THE IMAGE TO WEB AND SAVE AS PDF FROM THERE**

Finally, we want to determine 95 and 99.5% confidence intervals for each point.

To do this, we will refit our model with lme4 because the confint method with spaMM doesn't work well (keeps crashing). The model parameter estimates are identical.

```{r}

CI_mod <- lme4::glmer(Vacated ~ log(Median3) * Year_group + Status + (1|Sub_Area),
                  data = VacatedTerritories, family = binomial(logit))

CIs <- as.data.frame(cbind(confint(CI_mod, method = "Wald")[-1, ],
             confint(CI_mod, level = 0.995, method = "Wald")[-1, ])) %>%
  mutate(CI95 = paste(round(`2.5 %`, 4), round(`97.5 %`, 4), sep = "/"),
         CI995 = paste(round(`0.25 %`, 4), round(`99.75 %`, 4), sep = "/"),
         Estimate = format(as.numeric(spaMM::fixef(CI_mod))),
         Variable = names(spaMM::fixef(CI_mod))) %>% 
  dplyr::select(Variable, Estimate, CI95, CI995)

CI_25ha <- CIs %>% 
  dplyr::select(Variable, Estimate, CI95, CI995)

```

Determine R2 values for reviewers. To do this we need to use our lme4 model.

Add this to our CI estimates.

```{r}

R2 <- MuMIn::r.squaredGLMM(CI_mod)

CI_25ha %>% 
  mutate(R2m = round(R2[1, "R2m"], 2),
         R2c = round(R2[1, "R2c"], 2)) %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "bordered")

```
